---
title: "Integrative Analysis using MFA"
output:
  html_document:
    df_print: paged
---

# Introduction

Files `mRNA.csv` and `prots.csv` contain expression values from genes and proteins selected for being differentially expressed  from a TCGA breast cancer dataset.

```{r}
load("~/Dropbox (Nuevo Equipo VHIR10)/SotaCV/IODA-scripts/integrative_analysis_mixomics/data/expanded.matrix.genes.Rda")
matGenes <- expanded.matrix
dplyr::glimpse(matGenes)
tail(rownames(matGenes))
load("~/Dropbox (Nuevo Equipo VHIR10)/SotaCV/IODA-scripts/integrative_analysis_mixomics/data/expanded.matrix.prots.Rda")
matProts <- expanded.matrix
dplyr::glimpse(matProts)
tail(rownames(matProts))
```

# Applying MFA

## Data preparation

The expanded matrices contain either genes or proteins and a set of new variables derived from the selected annotations for each dataset.

```{r}
sum(colnames(matGenes)!=colnames(matProts))
numGOsInGenes <- sum(substr(rownames(matGenes),1,3) %in% "GO:")
numGOsInProts <- sum(substr(rownames(matProts),1,3) %in% "GO:")
numGenes <- nrow(matGenes) - numGOsInGenes
numProts <- nrow(matProts) - numGOsInProts
```

There are:

- `ncol(matGenes)` samples,
- `numGenes` genes,
- `numGOsInGenes` annotations in the _expanded gene matrix_,
- `numProts` proteins,
- `numGOsInProts` annotations in the _expanded protein matrix_.

In order to apply MFA we need a unique matrix with obtained by merging both datasets through its common dimension, which is set to be "rows", transposing the original matrices.


```{r}
onlyGenes <- t(matGenes)[,1:numGenes]
onlyProts <- t(matProts)[,1:numProts]
jointBasicData <- cbind(onlyGenes, onlyProts)
dplyr::glimpse(jointBasicData)
jointExpandedData <- cbind(t(matGenes), t(matProts))
dplyr::glimpse(jointExpandedData)
```
## Applying MFA

```{r}
library(FactoMineR)
grupsBasic <- c(numGenes,numProts)
names(grupsBasic) <- c("genes", "prots")
grupsExpanded <- c(numGenes,numGOsInGenes,numProts, numGOsInProts)
names(grupsExpanded) <- c("genes", "geneAnots", "prots", "protAnots")
```

```{r}
resMFAbasic<- MFA (base=jointBasicData, group=grupsBasic)
```


```{r}
resMFAexpanded<- MFA (base=jointExpandedData, group=grupsExpanded)
```

