---
title: "Integrative Analysis using MFA (v2. using 5 data sets after expansion)"
author: "Alex S&aacute;nchez-Pla and Ferran Brians&oacute;"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    theme: default
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

# Introduction

Files `exclusive.matrix.genes.Rda` and `exclusive.matrix.prots.Rda` contain expression values from genes and proteins selected for being differentially expressed  from a TCGA breast cancer dataset, plus (in additional rows) the new features coming from the respective annotation processes already performed with `annotate_and_expand` script.
File `intersect.matrix.Rda` contains only new features coming from the biological annotations in common between both data sources.

```{r}
load("../annotate_and_expand/results/input.matrix.genes.Rda", verbose = T)
matGenes <- my.matrix
dplyr::glimpse(matGenes)
tail(rownames(matGenes))

load("../annotate_and_expand/results/input.matrix.prots.Rda", verbose = T)
matProts <- my.matrix
dplyr::glimpse(matProts)
tail(rownames(matProts))

load("../annotate_and_expand/results/exclusive.matrix.genes.Rda", verbose = T)
annotGenes <- exclusive_expmat
dplyr::glimpse(annotGenes)
tail(rownames(annotGenes))

load("../annotate_and_expand/results/exclusive.matrix.prots.Rda", verbose = T)
annotProts <- exclusive_expmat
dplyr::glimpse(annotProts)
tail(rownames(annotProts))

load("../annotate_and_expand/results/intersect.matrix.Rda", verbose = T)
annotCommon <- intersect.matrix
dplyr::glimpse(annotCommon)
tail(rownames(annotCommon))

```


# Breast cancer subtypes 
Source [https://www.breastcancer.org/types/molecular-subtypes#section-luminal-a-breast-cancer]

Triple-negative or *basal-like* breast cancer is estrogen receptor-negative, progesterone receptor-negative, and HER2-negative. Triple-negative breast cancer is considered more aggressive than either luminal A or luminal B breast cancer.

*HER2-enriched* breast cancer is estrogen receptor-negative and progesterone receptor-negative and HER2-positive. HER2-enriched cancers tend to grow faster than luminal cancers and can have a worse prognosis, but are usually successfully treated with targeted therapy medicines aimed at the HER2 protein.

*Luminal A* breast cancer is estrogen receptor-positive and progesterone receptor-positive, HER2-negative, and has low levels of the protein Ki-67, which helps control how fast cancer cells grow. Luminal A cancers tend grow more slowly than other cancers, be lower grade, and have a good prognosis.

To get the cancer subtypes: 
```{r}
library(mixOmics)
data("breast.TCGA")
brca.subtype <- breast.TCGA$data.train$subtype
brca.subtype
table(brca.subtype)
```


# Data loading

The loaded matrices contain numeric values for either genes, proteins or a set of new variables (GO or similar annotations) derived from the selected annotations from each data set or the intersection between them.

```{r}
sum(colnames(matGenes)!=colnames(matProts))
sum(colnames(matProts)!=colnames(annotGenes))
sum(colnames(annotGenes)!=colnames(annotProts))
sum(colnames(annotProts)!=colnames(annotCommon))
#colnames(annotCommon)

numSamples <- ncol(annotGenes)
numGenes <- nrow(matGenes)
numProts <- nrow(matProts)
numGOsInGenes <- nrow(annotGenes)
numGOsInProts <- nrow(annotProts)
numGOsInCommon <- nrow(annotCommon)

```

There are:

- `r numSamples` samples,
- `r numGenes` genes,
- `r numGOsInGenes` annotation-derived features in the _exclusive annotations from gene matrix_,
- `r numProts` proteins,
- `r numGOsInProts` annotation-derived features in the _exclusive annotations from prots matrix_,
- `r numGOsInCommon` annotation-derived features from the _intersect annotations from both sources_.


# Preparing MFA

While the application of MFA is that of an exploratory analysis, so that there is no unique approach the list provided by [Husson et Josse (2010)](https://www.r-project.org/conferences/useR-2010/tutorials/Husson+Josse.pdf) is a useful "checklist" to organize the analysis:

0. Define the structure of the data (groups of variables and/or individuals)
1. Define groups of active and supplementary variables
2. Scale or not the variables
3. Perform MFA
4. Choose the number of dimensions to interpret
5. Simultaneously interpret the individuals and variables graphs
6. Study the groups of variables
7. Study the partial representations
8. Use indicators to enrich the interpretation


In order to apply MFA we need a unique matrix with obtained by merging all datasets through its common dimension, which is set to be "rows", transposing the original matrices.


```{r}
onlyGenes <- t(matGenes)#[,1:numGenes]
onlyProts <- t(matProts)#[,1:numProts]

jointBasicData <- cbind(onlyGenes, 
                        onlyProts)
jointBasicDataScaled <- cbind(scale(onlyGenes), 
                              scale(onlyProts))
dplyr::glimpse(jointBasicData)

jointExpandedData <- cbind(onlyGenes, 
                           onlyProts,
                           t(annotGenes),
                           t(annotProts),
                           t(annotCommon))
jointExpandedDataScaled <- cbind(scale(onlyGenes), 
                                 scale(onlyProts),
                                 scale(t(annotGenes)),
                                 scale(t(annotProts)),
                                 scale(t(annotCommon))
                                 )
dplyr::glimpse(jointExpandedData)
```

Groups are defined implicitly by the (number of) columns of the data matrix they are made of.


```{r}
numSamples
numGenes
numProts
numGOsInGenes
numGOsInProts
numGOsInCommon

grupsBasic <- c(numGenes,numProts)
grupsBasic
grupsExpanded <- c(numGenes,numProts,numGOsInGenes,numGOsInProts,numGOsInCommon)
grupsExpanded
```

Each group is given a name and it may be given a data type.

```{r}
basicNames <- names(grupsBasic) <- c("genes", "prots")
expandedNames <- names(grupsExpanded) <- c("genes", "prots", "geneAnots", "protAnots", "commonAnots")
basicTypes <- c("c", "c")
expandedTypes <-c(rep("c", 5))
```

## Defining active/suplementary data

Variables describing biological knowledge are linear combinations of one or more omics variable types so it seems reasonable to start considering these as supplementary.

Supplementary groups are described by the index of the group in the vector that defines how many variables in each groups.

```{r}
supplGroups <- c(3,4,5)
```

## Scaling the variables

A quick check at the numerical part of the dataset can suggest if the data needs to be centered/scaled.

```{r}
pal <- c(rep("blue",numGenes), rep("red", numProts))
boxplot(jointBasicData, col=pal, cex.axis=0.6, las=2)
boxplot(jointBasicDataScaled, col=pal, cex.axis=0.6, las=2)
```

```{r}
pal <- c(rep("blue",numGenes), rep("red", numProts), 
         rep("cyan", numGOsInGenes), rep("orange", numGOsInProts), 
         rep("darkmagenta", numGOsInCommon))
boxplot(jointExpandedData, col=pal, cex.axis=0.5, las=2)
boxplot(jointExpandedDataScaled, col=pal, cex.axis=0.4, las=2)
```


# Applying MFA to basic data

MFA is applied on the "basic" matrix formed only by expressions, that has been previously scaled.

```{r}
require(FactoMineR)
res.mfa.b <- MFA (base=jointBasicDataScaled, 
                      group=grupsBasic, 
                      name.group = names(grupsBasic),
                      type=basicTypes,
                      graph=FALSE)
```

```{r, eval=FALSE}
dimdesc(res.mfa.b)
```


## Representation of Groups

```{r}
plot(res.mfa.b,choix="group", palette=c("red", "blue"))
```

## Study of individuals and variables

```{r}
plot(res.mfa.b, choix="var",
     #invisible="quanti.sup",
     hab="group",
     lab.var=TRUE, lim.cos2.var = 0.5, 
     palette=c("black","blue", "red"))
```


```{r}
plot(res.mfa.b,choix="ind",partial="all",habillage="group",palette=c("black","blue", "red"))
```


```{r}
plot(res.mfa.b,choix="axes",habillage="group",palette=c("black","blue", "red"))
```


## Indivuals analysis with factoextra

Habillage parameter sets the phenotypes (= cancer subtypes)


```{r}
library(factoextra)

# Variables
#fviz_mfa_var(res.mfa.b, palette = ...)

# Individuals
pal <- c("green", "coral", "skyblue")
fviz_mfa_ind(res.mfa.b, habillage = brca.subtype, palette = pal) # color by cancer subtype
```


```{r}
# Individuals with all vars
pal <- c("green", "blue", "coral", "skyblue", "red")
fviz_mfa(res.mfa.b, habillage = brca.subtype, palette = pal)# color by groups
```



# Applying MFA to expanded data

MFA is applied on the "expanded" matrix formed by expressions and annotations, that has been previously scaled.

```{r}
require(FactoMineR)
res.mfa.x<- MFA (base=jointExpandedDataScaled, 
                      group=grupsExpanded, 
                      name.group = names(grupsExpanded),
                      type=expandedTypes,
                      num.group.sup = supplGroups, graph=FALSE)
```

```{r, eval=FALSE}
dimdesc(res.mfa.x)
```


## Representation of Groups

```{r}
pal <- c("darkmagenta", "blue", "red", "cyan", "orange")
plot(res.mfa.x,choix="group", palette=pal)
```

## Study of individuals and variables

```{r}
pal <- c("darkmagenta", "blue", "red", "cyan", "orange")
plot(res.mfa.x, choix="var",
     #invisible="quanti.sup",
     hab="group",
     lab.var=TRUE, lim.cos2.var = 0.5, 
     palette=pal)
```


```{r}
plot(res.mfa.x,choix="ind",partial="all",habillage="group",palette=c("black", "blue", "red"))
```


```{r}
pal <- c("black", "blue", "red", "cyan", "orange", "darkmagenta")
plot(res.mfa.x,choix="axes",habillage="group",palette=pal)
```


## Indivuals analysis with factoextra

Habillage parameter sets the phenotypes (= cancer subtypes)


```{r}
library(factoextra)

# Variables
#fviz_mfa_var(res.mfa.x, palette = ...)

# Individuals
pal <- c("green", "coral", "skyblue")
fviz_mfa_ind(res.mfa.x, habillage = brca.subtype, palette = pal) # color by cancer subtype
```


```{r}
# Individuals with all vars
pal <- c("green", "blue", "coral", "skyblue", "red")
fviz_mfa(res.mfa.x, habillage = brca.subtype, palette =  pal)# color by groups

```

# MFA redefining active/supplem. data

Now, we consider the group of features created from common annotations between genes and proteins also as principal data to the MFA analysis.

As said before, supplementary groups are described by the index of the group in the vector that defines how many variables in each groups.
In this case, we remove group 5 from the supplementary list.

```{r}
new.supplGroups <- c(3,4)
```


MFA is applied on the "expanded" matrix formed by expressions and annotations, that has been previously scaled.

```{r}
require(FactoMineR)
res.mfa.x2 <- MFA (base=jointExpandedDataScaled, 
                      group=grupsExpanded, 
                      name.group = names(grupsExpanded),
                      type=expandedTypes,
                      num.group.sup = new.supplGroups, graph=FALSE)
```

```{r, eval=FALSE}
dimdesc(res.mfa.x2)
```


## Representation of Groups

```{r}
pal <- c("orange", "blue", "red", "darkmagenta", "cyan")
plot(res.mfa.x2,choix="group", palette=pal)
```

## Study of individuals and variables

```{r}
pal <- c("orange", "blue", "red", "darkmagenta", "cyan")
plot(res.mfa.x2, choix="var",
     #invisible="quanti.sup",
     hab="group",
     lab.var=TRUE, lim.cos2.var = 0.5, 
     palette=pal)
```

Common annotations

GO:0019221 = cytokine-mediated signaling pathway
GO:0007165 = signal transduction
GO:0006357 = regulation of transcription by RNA polymerase II
GO:0045944 = positive regulation of transcription by RNA polymerase II

Gene annotations

GO:0006977 = DNA damage response, signal transduction by p53 class mediator resulting in cell cycle arrest
GO:0032355 = response to estradiol


```{r}
pal <- c("black", "blue", "red", "darkmagenta")
plot(res.mfa.x2,choix="ind",partial="all",habillage="group",palette=pal)
```


```{r}
pal <- c("black", "blue", "red", "darkmagenta", "cyan", "orange")
plot(res.mfa.x2,choix="axes",habillage="group",palette=pal)
```


## Indivuals analysis with factoextra

Habillage parameter sets the phenotypes (= cancer subtypes)


```{r}
library(factoextra)

# Variables
#fviz_mfa_var(res.mfa.x, palette = ...)

# Individuals
pal <- c("green", "coral", "skyblue")
fviz_mfa_ind(res.mfa.x2, habillage = brca.subtype, palette = pal) # color by cancer subtype
```


```{r}
# Individuals with all vars
pal <- c("green", "darkmagenta", "blue", "coral", "skyblue", "red")
fviz_mfa(res.mfa.x2, habillage = brca.subtype, palette =  pal)# color by groups

```
