---
title: "Report with results from the full analysis script (v.1)"
author: Ferran Brianso
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: TRUE
    theme: united
---

Targets' report corresponding to the functions that perform the full analysis, consisting in two parts:

1. Annotate any original expression matrix (with gene symbols as ids in rows, and sample ids in columns) and expands the annotated matrix to include the average (or sum) values for the annotations (as new extra rows).

2. Perform basic MFA analysis (as in "integrative_Analysis_MFA_v2.Rmd") ...


```{r libraries, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(heatmaply)
library(targets)
library(stats)
library(dplyr)
```

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 110)
```


### Raw data (before data prep) 

Data loaded from files **`r p.inFile1`** and **`r p.inFile2`**

Samples in rows; Features in columns 

```{r out.width = '95%', echo=FALSE}
show(tar_read(in_data1))
show(tar_read(in_data2))
```


### Data frame (before annotation) 

Input from **`r p.inFile1`**:
Features in ROWS: **`r nrow(tar_read(dframe1))`**
Samples in COLUMNS: **`r ncol(tar_read(dframe1))`**

Input from **`r p.inFile2`**:
Features in ROWS: **`r nrow(tar_read(dframe2))`**
Samples in COLUMNS: **`r ncol(tar_read(dframe2))`**


(Showing only partial output)
```{r out.width = '95%'}
tar_read(dframe1)[1:10,1:10]
tar_read(dframe2)[1:10,1:10]
```


### Heatmaps (before annotation)

Features in ROWS; Samples in COLUMNS

<!-- (Showing only partial output) -->
<!-- ```{r out.width = '95%'} -->
<!-- heatmap(data.matrix(dframe[1:40, 1:16]), -->
<!--         cexRow = 0.7, cexCol = 1, -->
<!--         Colv = NA) -->
<!-- ``` -->


#### Heatmaply
```{r Heatmap.Pre1, out.width = '95%'}
p <- heatmaply(data.matrix(dframe1), 
        #dendrogram = "row",
        xlab = "", ylab = "", 
        main = "",
        scale = "none",
        margins = c(60,100,40,20),
        #grid_color = "white",
        #grid_width = 0,
        titleX = FALSE,
        hide_colorbar = FALSE,
        branches_lwd = 0.1,
        label_names = c("Feature", "Sample", "Value"),
        fontsize_row = 5, fontsize_col = 5,
        labCol = colnames(data.matrix(dframe1)),
        labRow = rownames(data.matrix(dframe1)),
        heatmap_layers = theme(axis.line=element_blank())
        )
p
# save the widget
# library(htmlwidgets)
# saveWidget(p, file= "~/.../heatmap.html")
```

```{r Heatmap.Pre2, out.width = '95%'}
p <- heatmaply(data.matrix(dframe2), 
        #dendrogram = "row",
        xlab = "", ylab = "", 
        main = "",
        scale = "none",
        margins = c(60,100,40,20),
        #grid_color = "white",
        #grid_width = 0,
        titleX = FALSE,
        hide_colorbar = FALSE,
        branches_lwd = 0.1,
        label_names = c("Feature", "Sample", "Value"),
        fontsize_row = 5, fontsize_col = 5,
        labCol = colnames(data.matrix(dframe2)),
        labRow = rownames(data.matrix(dframe2)),
        heatmap_layers = theme(axis.line=element_blank())
        )
p
# save the widget
# library(htmlwidgets)
# saveWidget(p, file= "~/.../heatmap.html")
```


### Annotated Matrix

Gene Ontology used: **`r p.GO.onto`**

Min. number of genes required to pass the filter: **`r p.GO.miN`**

Annotated categories: **`r ncol(tar_read(categ_matrix1))`** (for *`r p.inFile1`*)

Annotated categories: **`r ncol(tar_read(categ_matrix2))`** (for *`r p.inFile2`*)

Shared annotated categories: **`r intersect(colnames(categ_matrix1), colnames(categ_matrix2))`**


(Showing only partial output)
```{r out.width = '95%'}
tar_read(categ_sums1)
tar_read(categ_sums2)
head(tar_read(categ_intersect1))
head(tar_read(categ_intersect2))
tar_read(categ_intersect)
```

Features in ROWS: **`r nrow(tar_read(annot_matrix1))`** (for *`r p.inFile1`*)

Features in ROWS: **`r nrow(tar_read(annot_matrix2))`** (for *`r p.inFile2`*)

Samples+Categs in COLUMNS: **`r ncol(tar_read(annot_matrix1))`** (for *`r p.inFile1`*)

Samples+Categs in COLUMNS: **`r ncol(tar_read(annot_matrix2))`** (for *`r p.inFile2`*)


(Showing only partial output)
```{r out.width = '95%'}
head(tar_read(annot_matrix1))
head(tar_read(annot_matrix2))
```

```{r Heatmap.Categs1, out.width = '95%', eval = FALSE}
p <- heatmaply(categ_matrix1, 
        dendrogram = "none",
        xlab = "", ylab = "", 
        main = "",
        scale = "none",
        margins = c(60,100,40,20),
        #grid_color = "white",
        #grid_width = 0,
        titleX = FALSE,
        hide_colorbar = FALSE,
        branches_lwd = 0.1,
        label_names = c("Feature", "Sample", "Value"),
        fontsize_row = 5, fontsize_col = 5,
        labCol = colnames(categ_matrix1),
        labRow = rownames(categ_matrix1),
        heatmap_layers = theme(axis.line=element_blank())
        )
p
# save the widget
# library(htmlwidgets)
# saveWidget(p, file= "~/.../heatmap.html")
```

```{r Heatmap.Categs2, out.width = '95%', eval = FALSE}
p <- heatmaply(categ_matrix2, 
        dendrogram = "none",
        xlab = "", ylab = "", 
        main = "",
        scale = "none",
        margins = c(60,100,40,20),
        #grid_color = "white",
        #grid_width = 0,
        titleX = FALSE,
        hide_colorbar = FALSE,
        branches_lwd = 0.1,
        label_names = c("Feature", "Sample", "Value"),
        fontsize_row = 5, fontsize_col = 5,
        labCol = colnames(categ_matrix2),
        labRow = rownames(categ_matrix2),
        heatmap_layers = theme(axis.line=element_blank())
        )
p
# save the widget
# library(htmlwidgets)
# saveWidget(p, file= "~/.../heatmap.html")
```


### Expanded Matrices

Gene Ontology used: **`r p.GO.onto`**

Min. number of genes applied to filter: **`r p.GO.miN`**

Annotated categories: **`r ncol(tar_read(categ_matrix1))`** (for *`r p.inFile1`*)

Annotated categories: **`r ncol(tar_read(categ_matrix2))`** (for *`r p.inFile2`*)

Features+Categs in ROWS: **`r nrow(tar_read(expd_matrix1))`** (for *`r p.inFile1`*)

Features+Categs in ROWS: **`r nrow(tar_read(expd_matrix2))`** (for *`r p.inFile2`*)

Samples in COLUMNS: **`r ncol(tar_read(expd_matrix1))`** (for *`r p.inFile1`*)

Samples in COLUMNS: **`r ncol(tar_read(expd_matrix2))`** (for *`r p.inFile2`*)



(Showing only partial output)
```{r out.width = '95%'}
tail(tar_read(expd_matrix1), ncol(categ_matrix1))[,1:6]
tail(tar_read(exc_exp_mat1), ncol(categ_matrix1))[,1:6]
dim(tar_read(exc_exp_mat1))
tail(tar_read(expd_matrix2), ncol(categ_matrix2))[,1:6]
tail(tar_read(exc_exp_mat2), ncol(categ_matrix2))[,1:6]
dim(tar_read(exc_exp_mat2))
tail(tar_read(int_exp_mat), 10)[,1:6]
dim(tar_read(int_exp_mat))
```


### Heatmap (of intersect matrix, after expansion)

Features+Categs in ROWS; Samples in COLUMNS

```{r Heatmap.Inter, out.width = '95%'}
p <- heatmaply(int_exp_mat, 
        #dendrogram = "row",
        xlab = "", ylab = "", 
        main = "",
        scale = "none",
        margins = c(60,100,40,20),
        #grid_color = "white",
        #grid_width = 0,
        titleX = FALSE,
        hide_colorbar = FALSE,
        branches_lwd = 0.1,
        label_names = c("Feature", "Sample", "Value"),
        fontsize_row = 5, fontsize_col = 5,
        labCol = colnames(int_exp_mat),
        labRow = rownames(int_exp_mat),
        heatmap_layers = theme(axis.line=element_blank())
        )
p
# save the widget
# library(htmlwidgets)
# saveWidget(p, file= "~/.../heatmap.html")
```


### Heatmap (after row expansion) 

Features+Categs in ROWS; Samples in COLUMNS

```{r Heatmap.Post1, out.width = '95%'}
p <- heatmaply(expd_matrix1, 
        #dendrogram = "row",
        xlab = "", ylab = "", 
        main = "",
        scale = "none",
        margins = c(60,100,40,20),
        #grid_color = "white",
        #grid_width = 0,
        titleX = FALSE,
        hide_colorbar = FALSE,
        branches_lwd = 0.1,
        label_names = c("Feature", "Sample", "Value"),
        fontsize_row = 5, fontsize_col = 5,
        labCol = colnames(expd_matrix1),
        labRow = rownames(expd_matrix1),
        heatmap_layers = theme(axis.line=element_blank())
        )
p
# save the widget
# library(htmlwidgets)
# saveWidget(p, file= "~/.../heatmap.html")
```

```{r Heatmap.Post2, out.width = '95%'}
p <- heatmaply(expd_matrix2, 
        #dendrogram = "row",
        xlab = "", ylab = "", 
        main = "",
        scale = "none",
        margins = c(60,100,40,20),
        #grid_color = "white",
        #grid_width = 0,
        titleX = FALSE,
        hide_colorbar = FALSE,
        branches_lwd = 0.1,
        label_names = c("Feature", "Sample", "Value"),
        fontsize_row = 5, fontsize_col = 5,
        labCol = colnames(expd_matrix2),
        labRow = rownames(expd_matrix2),
        heatmap_layers = theme(axis.line=element_blank())
        )
p
# save the widget
# library(htmlwidgets)
# saveWidget(p, file= "~/.../heatmap.html")
```


### MFA Data prep

NumSamples: **`r ncol(tar_read(dframe1))`**

NumGenes:  **`r nrow(tar_read(dframe1))`**

NumProts:  **`r nrow(tar_read(dframe2))`**

numGOsInGenes: **`r nrow(tar_read(exc_exp_mat1))`**

numGOsInProts: **`r nrow(tar_read(exc_exp_mat2))`**

numGOsInCommon: **`r nrow(tar_read(int_exp_mat))`**


```{r out.width = '95%', echo=FALSE}
show(tar_read(blocksBasic))
show(tar_read(blocksExp))
show(tar_read(groupInfo))
table(as.matrix(tar_read(groupInfo)))
```

supplementary groups: **`r p.supplGroupsMFA`**

```{r out.width = '95%', echo=FALSE}
#show(tar_read(expandedNames))
show(p.BlockNamesMFA)
#show(tar_read(expandedTypes))
show(p.BlockTypesMFA)
```


### MFA with Basic Blocks

```{r out.width = '95%', echo=FALSE}
# show(tar.read(res.mfa.b))
show(tar_read(res.mfa.b))
#dimdesc(tar_read(res.mfa.b))
```


Representation of Groups

```{r out.width = '95%', echo=FALSE}
plot(tar_read(res.mfa.b),choix="group", 
     palette=c("red", "blue"))
```


Study of individuals and variables

```{r out.width = '95%', echo=FALSE}
plot(tar_read(res.mfa.b), choix="var",
              #invisible="quanti.sup",
              hab="group",
              lab.var=TRUE, lim.cos2.var = 0.5, 
              palette=c("black", "blue", "red"))

plot(tar_read(res.mfa.b), choix="ind",
     partial="all",
     habillage="group",
     palette=c("black","blue", "red"))

plot(tar_read(res.mfa.b), choix="axes", 
     habillage="group", 
     palette=c("black","blue", "red"))
```


Indivuals analysis with factoextra

```{r out.width = '95%', echo=FALSE}
# Habillage parameter sets the phenotypes (= cancer subtypes)
fviz_mfa_ind(tar_read(res.mfa.b), habillage = tar_read(groupInfo), 
             palette = c("green", "coral", "skyblue")) # color by cancer subtype
```

Individuals analysis with all vars

```{r out.width = '95%', echo=FALSE}
# Habillage parameter sets the phenotypes (= cancer subtypes)
fviz_mfa(tar_read(res.mfa.b), habillage = tar_read(groupInfo), 
         palette = c("green", "blue", "coral", "skyblue", "red"))# color by groups
```


### MFA with Expanded Data

MFA is applied on the "expanded" matrix formed by expressions and annotations, that has been previously scaled.

```{r out.width = '95%', echo=FALSE}
# show(tar.read(res.mfa.x))
show(tar_read(res.mfa.x))
#dimdesc(tar_read(res.mfa.x))
```

Representation of Groups

```{r out.width = '95%', echo=FALSE}
plot(tar_read(res.mfa.x), choix="group", 
     palette=c("darkmagenta", "blue", "red", "cyan", "orange"))
```

Study of individuals and variables

```{r out.width = '95%', echo=FALSE}
plot(tar_read(res.mfa.x), choix="var",
              #invisible="quanti.sup",
              hab="group",
              lab.var=TRUE, lim.cos2.var = 0.5, 
              palette=c("darkmagenta", "blue", "red", "cyan", "orange"))

plot(tar_read(res.mfa.x), choix="ind",
     partial="all",
     habillage="group",
     palette=c("black", "blue", "red"))

plot(tar_read(res.mfa.x), choix="axes", 
     habillage="group", 
     palette=c("black", "blue", "red", "cyan", "orange", "darkmagenta"))
```

Indivuals analysis with factoextra

```{r out.width = '95%', echo=FALSE}
# Habillage parameter sets the phenotypes (= cancer subtypes)
fviz_mfa_ind(tar_read(res.mfa.x), habillage = tar_read(groupInfo), 
             palette = c("green", "coral", "skyblue")) # color by cancer subtype
```

Individuals analysis with all vars

```{r out.width = '95%', echo=FALSE}
# Habillage parameter sets the phenotypes (= cancer subtypes)
fviz_mfa(tar_read(res.mfa.x), habillage = tar_read(groupInfo), 
         palette = c("green", "blue", "coral", "skyblue", "red"))# color by groups
```


### MFA redefining active/supplem. data

Now, we consider the group of features created from common annotations between genes and proteins also as principal data to the MFA analysis.

As said before, supplementary groups are described by the index of the group in the vector that defines how many variables in each groups.

In this case, we remove group 5 from the supplementary list.

MFA is applied on the "expanded" matrix formed by expressions and annotations, that has been previously scaled.

```{r out.width = '95%', echo=FALSE}
# show(tar.read(res.mfa.x))
show(tar_read(res.mfa.x2))
#dimdesc(tar_read(res.mfa.x))
```

Representation of Groups

```{r out.width = '95%', echo=FALSE}
plot(tar_read(res.mfa.x2), choix="group", 
     palette=c("orange", "blue", "red", "darkmagenta", "cyan"))
```

Study of individuals and variables

```{r out.width = '95%', echo=FALSE}
plot(tar_read(res.mfa.x2), choix="var",
              #invisible="quanti.sup",
              hab="group",
              lab.var=TRUE, lim.cos2.var = 0.5, 
              palette=c("orange", "blue", "red", "darkmagenta", "cyan"))

plot(tar_read(res.mfa.x2), choix="ind",
     partial="all",
     habillage="group",
     palette=c("black", "blue", "red", "darkmagenta"))

plot(tar_read(res.mfa.x2), choix="axes", 
     habillage="group", 
     palette=c("black", "blue", "red", "darkmagenta", "cyan", "orange"))
```


Indivuals analysis with factoextra

```{r out.width = '95%', echo=FALSE}
# Habillage parameter sets the phenotypes (= cancer subtypes)
fviz_mfa_ind(tar_read(res.mfa.x2), habillage = tar_read(groupInfo), 
             palette = c("green", "coral", "skyblue")) # color by cancer subtype
```

Individuals analysis with all vars

```{r out.width = '95%', echo=FALSE}
# Habillage parameter sets the phenotypes (= cancer subtypes)
fviz_mfa(tar_read(res.mfa.x2), habillage = tar_read(groupInfo), 
         palette = c("green", "darkmagenta", "blue", "coral", "skyblue", "red"))# color by groups
```
