---
title: "Report with results from the 'Annotate and Expand Omics Data Matrix' script (v.1)"
author: Ferran Brianso
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: TRUE
    theme: united
---

Targets' report corresponding to the functions that annotate any original expression matrix (with gene symbols as ids in rows, and sample ids in columns) and expands the annotated matrix to include the average (or sum) values for the annotations (as new extra rows)

```{r libraries, echo=FALSE}
library(ggplot2)
library(heatmaply)
library(targets)
library(stats)
```

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 110)
```

### Raw data (before data prep) 

Data loaded from files **`r p.inFile1`** and **`r p.inFile2`**

Samples in rows; Features in columns 

```{r out.width = '95%', echo=FALSE}
show(tar_read(in_data1))
show(tar_read(in_data2))
```

### Data frame (before annotation) 

Input from **`r p.inFile1`**:
Features in ROWS: **`r nrow(tar_read(dframe1))`**
Samples in COLUMNS: **`r ncol(tar_read(dframe1))`**

Input from **`r p.inFile2`**:
Features in ROWS: **`r nrow(tar_read(dframe2))`**
Samples in COLUMNS: **`r ncol(tar_read(dframe2))`**


(Showing only partial output)
```{r out.width = '95%'}
tar_read(dframe1)[1:10,1:10]
tar_read(dframe2)[1:10,1:10]
```

### Heatmaps (before annotation)

Features in ROWS; Samples in COLUMNS

<!-- (Showing only partial output) -->
<!-- ```{r out.width = '95%'} -->
<!-- heatmap(data.matrix(dframe[1:40, 1:16]), -->
<!--         cexRow = 0.7, cexCol = 1, -->
<!--         Colv = NA) -->
<!-- ``` -->

#### Heatmaply
```{r Heatmap.Pre1, out.width = '95%'}
p <- heatmaply(data.matrix(dframe1), 
        #dendrogram = "row",
        xlab = "", ylab = "", 
        main = "",
        scale = "none",
        margins = c(60,100,40,20),
        #grid_color = "white",
        #grid_width = 0,
        titleX = FALSE,
        hide_colorbar = FALSE,
        branches_lwd = 0.1,
        label_names = c("Feature", "Sample", "Value"),
        fontsize_row = 5, fontsize_col = 5,
        labCol = colnames(data.matrix(dframe1)),
        labRow = rownames(data.matrix(dframe1)),
        heatmap_layers = theme(axis.line=element_blank())
        )
p
# save the widget
# library(htmlwidgets)
# saveWidget(p, file= "~/.../heatmap.html")
```

```{r Heatmap.Pre2, out.width = '95%'}
p <- heatmaply(data.matrix(dframe2), 
        #dendrogram = "row",
        xlab = "", ylab = "", 
        main = "",
        scale = "none",
        margins = c(60,100,40,20),
        #grid_color = "white",
        #grid_width = 0,
        titleX = FALSE,
        hide_colorbar = FALSE,
        branches_lwd = 0.1,
        label_names = c("Feature", "Sample", "Value"),
        fontsize_row = 5, fontsize_col = 5,
        labCol = colnames(data.matrix(dframe2)),
        labRow = rownames(data.matrix(dframe2)),
        heatmap_layers = theme(axis.line=element_blank())
        )
p
# save the widget
# library(htmlwidgets)
# saveWidget(p, file= "~/.../heatmap.html")
```

### Annotated Matrix

Gene Ontology used: **`r p.GO.onto`**

Min. number of genes required to pass the filter: **`r p.GO.miN`**

Annotated categories: **`r ncol(tar_read(categ_matrix1))`** (for *`r p.inFile1`*)

Annotated categories: **`r ncol(tar_read(categ_matrix2))`** (for *`r p.inFile2`*)

```{r out.width = '95%'}
tar_read(categ_sums1)
tar_read(categ_sums2)
```

Features in ROWS: **`r nrow(tar_read(annot_matrix1))`** (for *`r p.inFile1`*)

Features in ROWS: **`r nrow(tar_read(annot_matrix2))`** (for *`r p.inFile2`*)

Samples+Categs in COLUMNS: **`r ncol(tar_read(annot_matrix1))`** (for *`r p.inFile1`*)

Samples+Categs in COLUMNS: **`r ncol(tar_read(annot_matrix2))`** (for *`r p.inFile2`*)


(Showing only partial output)
```{r out.width = '95%'}
head(tar_read(annot_matrix1))
head(tar_read(annot_matrix2))
```

```{r Heatmap.Categs1, out.width = '95%', eval = FALSE}
p <- heatmaply(categ_matrix1, 
        dendrogram = "none",
        xlab = "", ylab = "", 
        main = "",
        scale = "none",
        margins = c(60,100,40,20),
        #grid_color = "white",
        #grid_width = 0,
        titleX = FALSE,
        hide_colorbar = FALSE,
        branches_lwd = 0.1,
        label_names = c("Feature", "Sample", "Value"),
        fontsize_row = 5, fontsize_col = 5,
        labCol = colnames(categ_matrix1),
        labRow = rownames(categ_matrix1),
        heatmap_layers = theme(axis.line=element_blank())
        )
p
# save the widget
# library(htmlwidgets)
# saveWidget(p, file= "~/.../heatmap.html")
```

```{r Heatmap.Categs2, out.width = '95%', eval = FALSE}
p <- heatmaply(categ_matrix2, 
        dendrogram = "none",
        xlab = "", ylab = "", 
        main = "",
        scale = "none",
        margins = c(60,100,40,20),
        #grid_color = "white",
        #grid_width = 0,
        titleX = FALSE,
        hide_colorbar = FALSE,
        branches_lwd = 0.1,
        label_names = c("Feature", "Sample", "Value"),
        fontsize_row = 5, fontsize_col = 5,
        labCol = colnames(categ_matrix2),
        labRow = rownames(categ_matrix2),
        heatmap_layers = theme(axis.line=element_blank())
        )
p
# save the widget
# library(htmlwidgets)
# saveWidget(p, file= "~/.../heatmap.html")
```


### Expanded Matrix

Gene Ontology used: **`r p.GO.onto`**

Min. number of genes applied to filter: **`r p.GO.miN`**

Annotated categories: **`r ncol(tar_read(categ_matrix1))`** (for *`r p.inFile1`*)

Annotated categories: **`r ncol(tar_read(categ_matrix2))`** (for *`r p.inFile2`*)

Features+Categs in ROWS: **`r nrow(tar_read(expd_matrix1))`** (for *`r p.inFile1`*)

Features+Categs in ROWS: **`r nrow(tar_read(expd_matrix2))`** (for *`r p.inFile2`*)

Samples in COLUMNS: **`r ncol(tar_read(expd_matrix1))`** (for *`r p.inFile1`*)

Samples in COLUMNS: **`r ncol(tar_read(expd_matrix2))`** (for *`r p.inFile2`*)


(Showing only partial output)
```{r out.width = '95%'}
tail(tar_read(expd_matrix1), ncol(categ_matrix1))[,1:10]
tail(tar_read(expd_matrix2), ncol(categ_matrix2))[,1:10]
```

### Heatmap (after row expansion) 

Features+Categs in ROWS; Samples in COLUMNS

```{r Heatmap.Post1, out.width = '95%'}
p <- heatmaply(expd_matrix1, 
        #dendrogram = "row",
        xlab = "", ylab = "", 
        main = "",
        scale = "none",
        margins = c(60,100,40,20),
        #grid_color = "white",
        #grid_width = 0,
        titleX = FALSE,
        hide_colorbar = FALSE,
        branches_lwd = 0.1,
        label_names = c("Feature", "Sample", "Value"),
        fontsize_row = 5, fontsize_col = 5,
        labCol = colnames(expd_matrix1),
        labRow = rownames(expd_matrix1),
        heatmap_layers = theme(axis.line=element_blank())
        )
p
# save the widget
# library(htmlwidgets)
# saveWidget(p, file= "~/.../heatmap.html")
```

```{r Heatmap.Post2, out.width = '95%'}
p <- heatmaply(expd_matrix2, 
        #dendrogram = "row",
        xlab = "", ylab = "", 
        main = "",
        scale = "none",
        margins = c(60,100,40,20),
        #grid_color = "white",
        #grid_width = 0,
        titleX = FALSE,
        hide_colorbar = FALSE,
        branches_lwd = 0.1,
        label_names = c("Feature", "Sample", "Value"),
        fontsize_row = 5, fontsize_col = 5,
        labCol = colnames(expd_matrix2),
        labRow = rownames(expd_matrix2),
        heatmap_layers = theme(axis.line=element_blank())
        )
p
# save the widget
# library(htmlwidgets)
# saveWidget(p, file= "~/.../heatmap.html")
```
